---
title: "ST 558 Project 3"
author: "Katelyn Settlemyre, Xi Zeng"
date: "2022-10-13"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Package List  
Below is the package list needed for this project.
```{r， warning = FALSE, message=FALSE}
library(tidyverse)
library(caret)
library(MASS)
library(randomForest)
```

# Introduction  
The data set used for analysis is the `OnlineNewsPopularity` data set from UCI Machine Learning Repository. The data set has 61 variables in total, as well as 39644 observations. First, we will take a subset data set by filter according to the data channel. The lifestyle channel is first chosen for analysis.  
For our project, we will take variables(With description of each variable):  
+ `n_tokens_title`: Number of words in the title  
+ `n_tokens_content`: Number of words in the content  
+ `num_imgs`: Number of images  
+ `average_token_length`: Average length of the words in the content  
+ `is_weekend`: Was the article published on the weekend?  
+ `global_rate_positive_words`: Rate of positive words in the content  
+ `num_keywords`: Number of keywords in the metadata  

The response variable that we are interested in is the `shares` variable, which stands for the number of sales in social network. A basic EDA will be performed at first to examine the relationship between covariates and `shares`. Then predicted modeling is applied to predict the number of sales, which will be illustrated in detail below.  

# Data read in
```{r}
#Read in data
news <- read_csv("OnlineNewsPopularity.csv", show_col_types = FALSE)

#Subset data with a single data channel
subnews <- subset(news, data_channel_is_lifestyle == 1)

#Select variables for modeling and EDA
subnews <- subnews %>% dplyr::select(n_tokens_title,n_tokens_content,num_imgs,average_token_length,is_weekend,global_rate_positive_words,shares,num_keywords)
subnews$is_weekend <- as.factor(subnews$is_weekend)
```
 
# Data Split  
In this part, our data set is split into training set and test set, with training set contains 70% of the data and test set contains the other 30%.  
```{r}
#Set seed for reproduction
set.seed(123)

#Split data
index <- createDataPartition(subnews$shares,p = 0.7,list = FALSE)
train <- subnews[index,]
test <- subnews[-index,]

```


# Summarizations and EDA  

In this part, basic summary statistics for the predictors are calculated. Also, some plots,including scatter plots, barplots, and boxplots, as well as frequency tables are generated to examine the relationship between the variables.  
```{r}
#Define function for producing summary statistics
sum <- function(x){
  min <- min(x)
  mean <- mean(x)
  median <- median(x)
  max <- max(x)
  sd <- sd(x)
  return(list(Minimum = min, Mean = mean, Median = median, Maximum = max, `Standard Deviation` = sd(x)))
}

#Calculate summary statistic for variables
sum(subnews$n_tokens_title)
sum(subnews$n_tokens_content)
```

We also show correlations between our selected variables. We show both a correlation matrix and correlation plot.
```{r}
#correlation matrix
(corrs <- cor(subnews[,-5]))

#correlation plot
corrplot::corrplot(corrs)
```
Our correlation matrix and plot shows that we don't have strong relationships between variables. The strongest relationship is a correlation of 0.46 between `n_tokens_content` and `num_imgs`.

After producing summary statistics, some plots and tables are shown below:  

## Frequency tables  

Below shows the frequency table of number of keywords.  
```{r}
 table(train$num_keywords)
```

According to the table, most of the articles have more than 5 keywords. We assume that more keywords in the metadata will increase shares.  

Here we have a frequency table of number of articles published on a weekend.
```{r}
table(subnews$is_weekend)
```
According to this table, we see that approximately one fifth of the articles in our data are published on the weekend.

## Scatter plots

Below are the scatter plots for the chosen variables.  
```{r}
ggplot(data = train, aes(x = global_rate_positive_words, y = shares))+
geom_point(aes(color = is_weekend)) +
  labs(x = "Rate of positive words in the content",
       y = "Number of Shares",
       title = "Scatter Plot of Shares vs Rate of positive words in the content") +
  scale_color_discrete(name = "Weekend Published", labels = c("No", "Yes"))
```

According to the scatter plot above, it seems that when the rate of positive words exceeds 0.075, the number of shares of the article is relatively small. Also, those articles with high shares are mostly published during weekdays rather than weekend.  

## Barplots

Below are the bar plot for the number of keywords：  

```{r}
# Create bar plot of predictor "is_weekend"
g <- ggplot(data = train, aes(x = num_keywords, fill= is_weekend))
g + geom_bar(aes(fill= is_weekend),position = "dodge") +
  xlab("Number of keywords")+
  ggtitle("Bar Plot of number of keyword")+
  scale_fill_discrete(name = "Weekend Published", labels = c("No", "Yes"))
```

According to the bar plot, the frequency count of the each number of keywords is shown, also ,it will be group into weekday and weekend to see if there is any difference between them.  

## Boxplots  

Below are the box plot for sales:  

```{r}
g <- ggplot(train, aes(x = is_weekend, y = shares))
g + geom_point(aes(color = is_weekend), position = "jitter")+
  ggtitle("Jitter Plot for shares on weekday/weekend") + xlab("is_weekend")+
  scale_color_discrete(name = "Weekend Published", labels = c("No", "Yes"))
```

A jitter plot is generated to see the spread of shares data in weekdays and weekends.  

# Model fitting  

Below is the model fitting part. In this part, four models are fitted. They are: linear regression model, random forest model and boosted tree model.

## Linear Regression  
We fit two different linear regression models here. Linear regression is a basic method to find a linear relationship between a response variable and one or more predictor variables. Here we will fit models using a forward selection of the predictor variables as well as a subset selection of variables and interaction terms.  
First we fit the forward selection.
```{r}
#Use forward selection to determine the predictors used for the model
lm_forward <- train(shares~ .,
                    data = train,
                    method = "glmStepAIC",
                    direction = "forward",
                    trControl = trainControl("cv",number=10),
                    trace = FALSE)
lm_forward$finalModel

#Fit model with selected variables on test data
lmod_pred <- predict(lm_forward, newdata = test)

#Calculate test error
lmod1_RMSE <- RMSE(lmod_pred,test$shares)
lmod1_RMSE
```

Now we fit the subset selection.
```{r}
#Fit linear model using subset method
lm_subset <- train(shares~.^2,
                   data = train,
                   method = "lmStepAIC",
                   trControl = trainControl("cv",number=5),
                   trace = FALSE)
lm_subset$finalModel
lm_subset$results

#fit with test data
lsub_pred <- predict(lm_subset, newdata = test)

#test error:
(lmod2_RMSE <- RMSE(lsub_pred,test$shares))
```


## Random Forest Model  

Here we fit a random forest model. A random forest is an method where multiple tree models are fit from bootstrap samples using a subset of predictor variables for each bootstrap sample. The final prediction is an average of the bootstrap predictions.

```{r}
#set tuning parameters
rand_grid <- data.frame(mtry=1:5)

#train model
rand_fit <- train(shares~.,
                  data = train,
                  method = "rf",
                  preProcess = c("center", "scale"),
                  trControl = trainControl(method = "repeatedcv", number = 5, repeats = 3),
                  tuneGrid = rand_grid)
rand_fit$bestTune
rand_fit$finalModel

#fit with test data
rand_pred <- predict(rand_fit, newdata = test)

#test error:
(rand_RMSE <- RMSE(rand_pred, test$shares))
```


## Boosted Tree Model  

Below is the process of fitting a boosted tree model. Model performance is tested by fitting the final tuned model on test set and calculate the test RMSE. Here, I use the

```{r}
#Set tuning grid for boosted tree model
boost_grid <- expand.grid(n.trees = c(25,50,100,150,200),
                          interaction.depth = c(1:5),
                          shrinkage = c(0.1,0.2,0.3,0.4,0.5),
                          n.minobsinnode = 10)
boost_grid

#Train the model
boost_fit <- train(shares ~., data = train,
                   method = "gbm",
                   trControl = trainControl(method = "repeatedcv", number = 5, repeats = 3),
                   preProcess = c("center", "scale"),
                   tuneGrid = boost_grid,
                   verbose = FALSE)

boost_fit$bestTune

#Fit the final boosted tree model with test data
boost_pred <- predict(boost_fit, newdata = test)

#Calculate test RMSE for model performance
boost_RMSE <- RMSE(boost_pred,test$shares)
boost_RMSE
```